---
title: "CourseProject-RegModels"
output: html_document
---

###Executive Summary

```{r echo=FALSE}
library(datasets); data(mtcars)
mtcars$cyl <- factor(mtcars$cyl);mtcars$gear <- factor(mtcars$gear)
mtcars$vs <- factor(mtcars$vs);mtcars$carb <- factor(mtcars$carb)
mtcars$am[mtcars$am =="0"] <- "automatic"
mtcars$am[mtcars$am =="1"] <- "manual"
mtcars$am <- as.factor(mtcars$am)
st<-t.test(mpg ~ am, data=mtcars)
t.test(mpg ~ am, data=mtcars)

```

An analysis was done on a data set of a collection of cars, We were interested in exploring the relationship between different varibales from a set of eleven variables, and mile per gallon(MPG). Dataset from the 1974 Motor Trend has been used to find the answers to these two questions:

    “Is an automatic or manual transmission better for MPG”
    "Quantify the MPG difference between automatic and manual transmissions"
    
By incorporating hypothesis testing and simple linear regression, it is determined that there is significant difference between the mean MPG for automatic and manual transmission cars, on average manual transmission cars have `r st$estimate[2]- st$estimate[1]` more MPG. Moreover, in order to regulate for other confounding variables such as weight, Number of cylinders and horsepower of the car, multivariate regression model is used to find out the impact of other variables on MPG.Simple and multivariate regression models were validated by using ANOVA test. Results from the multivariate regression disclose that on average manual transmission cars get 1.81 miles more per gallon as compare to automatic transmission cars.


###Exploratory Data Analysis


From the exploratory analysis (Figure 1,Appendix) we are exploring the relation between the mpg and its effects of car transmission type, we can observe from box-and-whisker plot (Figure 1,Appendix) that there is a steady increase in mpg when the transmission for the car used is manual, the same can be seen from the t.test.The p-value `r st$p.value` is very small which assured that there is actual distinction between the two groups. 95% confidence interval shows that mpg in manual cars can be lower than automatic cars and it varies from `r st$conf.int`. Hence the null Hypothesis is rejected.There exist a significant difference in the mean MPG between manual transmission cars and automatic transmission cars. 
In (Figure 2, Appendix) done through scatterplot of all the variables in the dataset we can observe that there is a significant correlation between mpg and the other variables of interest like cyl, disp, hp, draft, wt, vs and am. 


###Regression Analysis

####Simple Linear Regression Model

```{r echo=FALSE}
sfit <- lm(mpg~am,mtcars)
summary(sfit)
```

Null hypothesis is rejected and it can be said that, there exists a linear relation between the predictor variable mpg and am.Intercepts and Coefficient can be explained that, on average automatic transmission cars has `r sfit$coefficients[1]` MPG and manual transmission cars has `r sfit$coefficients[1] + sfit$coefficients[2]` ( `r sfit$coefficients[1]` + `r sfit$coefficients[2]` ).The value of R^2 is `r summary(sfit)$r.squared`, which means this model only explain `r round((summary(sfit)$r.squared) * 100)`% of the variance.

####Multivariate Regression Model

#####Best Regression Model

In order to choose the best model, we use the stepwise selection (forward, backward, both) using the stepAIC( ) function from the MASS package.

```{r echo=FALSE}
library(MASS)
fit <- lm(mpg~.,data=mtcars)
bestmodel <- stepAIC(fit, direction="both")
```

The following code chunk shows the best regression model that is obtained by the interactions of several variables. The best model consists of the variables, cyl, wt and hp as confounders and am as the independent variable. As we can see the R-square coefficient of `r summary(bestmodel)$r.squared` is the value we can obtain by fitting the best model.

#####Analysis of variances (ANOVA Test)

We now derive from the anova for different cases involving a. transmission, b. all variables and c. best fit variable combination of predictors cyl, hp, wt and am.
```{r echo=FALSE}
base <- lm(mpg ~ am, data = mtcars)
all <- lm(mpg~.,data=mtcars)
anova(all,base,bestmodel)
```

Note that the smaller values for the p-value in the anova above indicates that the confounding variables cyl, wt and hp are statistically insignificant; and we reject the hypothesis that these variables do not contribute to the accuracy of our model.

With the p-value 1.5e-05, null hypothesis is rejected and it can be claimed that the multivariate model of regression is significantly different from that of simple model.

#####Residual - Diagnostics Analysis

Please see the residual plots for the chosen regression model (Appendix). We can conclude the following from the plot:

    The points in the Residuals vs Fitted plot are randomly scattered with no particular pattern.
    The points in the Normal Q–Q plot are more-or-less on the line, indicating that the residuals follow a normal distribution.
    In both the Scale–Location plot and the Residuals vs Leverage plots, the points are in a group with none too far from the center.

After examing the plots, we can say that residuals are normally distributed and are homoskedastic. Now estimates from the multivariate model can be interpreted.

```{r echo=FALSE}
mfit <- lm(mpg~cyl + hp + wt + am,mtcars)
coef(mfit)
```

```{r echo=FALSE}
cfit <- lm(mpg~cyl,mtcars)
coef(cfit)
```

Multivariate regression model explain `r summary(bestmodel)$r.squared * 100`% variance. Furthermore it can be seen that wt and to some extent hp confound the relationship between am and mpg.On Average manual transmission cars has `r mfit$coefficients[6]` MPGs more than automatic transmission cars. It is concluded that, it is better for MPG to have manual transmission cars vs automatic cars.

###Conclusions

From our multiple regression analysis above, we conclude the following:

    The cars with a manual transmission are better for mpg than cars with an automatic transmission.
    The rate of change of the conditional mean mpg with respect to am is about `r mfit$coefficients[6]`, and we are 95% confident that this value varies between -1.06 and 4.68.
    If we increase the number of cylinders from 4 to 8, mpg will decrease by a factor of 3 concluding a direct relationship.


###Appendix

##### Exploratory Data Analysis
```{r echo=FALSE,fig.width = 4, fig.height = 4}
plot(mpg ~ am, mtcars,main="Figure 1. Boxplot of MPG by transmission type", xlab="Transmission mode", ylab = "MPG")
plot(mtcars,main="Figure 2. Correlation Matrix mtcars")
```

##### Regression Analysis Plots

#####Residual plots for the regression model

```{r echo=FALSE}
par(mfrow = c(2,2))
plot(bestmodel)
```

#####Cooks Distance plot for the regression model

```{r echo=FALSE}
par(mfrow = c(1,2))
plot(bestmodel, which=c(4,6))
```

####Diagnostics

We want to identify the impact of an observation on the regression coefficients, and one approach is to consider how much the regression coefficient values change if the observation was not considered.

```{r echo=FALSE}
leverage<-hatvalues(bestmodel)
head(sort(leverage,decreasing=TRUE),3)
```

Please see the appendix section for the cooks distance plot for the chosen regression model.

```{r echo=FALSE}
tail(sort(cooks.distance(bestmodel)),3)
```

From the cooks distance plot above we can confirm our analysis was correct, as the same cars are mentioned in the residual plots.
